Local real space operators for translating Stokes Q/U to E/B maps on the sphere

### 2017-09-20 article/sections/numerical_implementation.tex
updates to the notebook evaluating the high res analysis, added latitude position of the mask as a input variable

At first glance, the above remarks seem to hold true only for the E mode spectra, since the errors in the B mode spectra seem to be significantly larger as seen in the middle row of \fig{fig:eb-spectra_rad_cutoff}. We understand that the higher error in the B mode spectra is primarily a combined effect of the low amplitude of the B mode spectrum as compared to the E mode spectrum and limited numerical accuracy of our simple convolution algorithm. We ran a number of tests to confirm these as reasons for the observed discrepant behaviour. 
%
\begin{figure}[!t] 
\centering
%\vspace{-5cm}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}} \\[-2ex]
\caption{\textit{Top:} The figures on left and right depict the E mode spectra and the B mode spectra respectively. \textit{Middle:} These figures depict the relative percentage difference between the spectra derived from full sky, locally evaluated E \& B  maps and the reference spectra. \textit{Bottom:} These figures depict the same results as shown in the middle row, except that the regions with $|b|>30^\circ$ of the maps have been masked for this analysis. 
The colors indicate the spectra derived from imposing different radial cutoff as indicated by the legends. The gray bands indicate the cosmic variance scatter around the fiducial spectra. The dashed horizontal black lines mark the 2\% error.}
\label{fig:eb-spectra_rad_cutoff}
\end{figure}
%
We exchanged the E and B mode spectra, \revisit{(taking care to set $C_{\ell}^{TE}=0$)} and noted that the trends observed in the error plots are reversed, clearly indicating a dependence on the amplitude of the spectrum. We noted that the polar caps were particularly error prone, as seen in \fig{fig:eb-maps-compare}. We excised these error prone regions, specifically masking portions of the maps with $|b|>30^{\circ}$ to find that the errors on the B mode spectra revert to having very similar structure as that of the E mode spectra, as seen in the bottom row of \fig{fig:eb-spectra_rad_cutoff}. Finally we modify the use of standard Healpix routines to mimic the local convolution E \& B construction (see \app{sec:convolution_err} for details). The results from this exercise also concur with our understanding of the discrepant behaviour for the B-mode spectrum.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Separating Stokes Q \& U maps corresponding to E \& B modes of polarization}
Here 
%
\begin{figure}[!h] 
\centering
\subfigure[E Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-healpix.pdf}}
\subfigure[E Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-2beta.pdf}}
\subfigure[E Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-diff.pdf}}
\subfigure[E U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-healpix.pdf}}
\subfigure[E U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-2beta.pdf}}
\subfigure[E U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-diff.pdf}}

@@

%
\begin{figure}[!h] 
\centering
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{}
\label{fig:equ-bqu-spectra_rad_cutoff}
\end{figure}
%
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
%--------------------------------------------------------
%--------------------------------------------------------

>>>>

At first glance, the above remarks seem to hold true only for the E mode spectra, since the errors in the B mode spectra seem to be significantly larger as seen in the middle row of \fig{fig:eb-spectra_rad_cutoff}. We understand that the higher error in the B mode spectra is primarily a combined effect of the low amplitude of the B mode spectrum as compared to the E mode spectrum and limited numerical accuracy of our simple convolution algorithm. We ran a number of tests to confirm these as reasons for the observed discrepant behaviour. 
%
\begin{figure}[!t] 
\centering
%\vspace{-5cm}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}} \\[-2ex]
\caption{\textit{Top:} The figures on left and right depict the E mode spectra and the B mode spectra respectively. \textit{Middle:} These figures depict the relative percentage difference between the spectra derived from full sky, locally evaluated E \& B  maps and the reference spectra. \textit{Bottom:} These figures depict the same results as shown in the middle row, except that the regions with $|b|>30^\circ$ of the maps have been masked for this analysis. 
The colors indicate the spectra derived from imposing different radial cutoff as indicated by the legends. The gray bands indicate the cosmic variance scatter around the fiducial spectra. The dashed horizontal black lines mark the 2\% error.}
\label{fig:eb-spectra_rad_cutoff}
\end{figure}
%
We exchanged the E and B mode spectra, \revisit{(taking care to set $C_{\ell}^{TE}=0$)} and noted that the trends observed in the error plots are reversed, clearly indicating a dependence on the amplitude of the spectrum. We noted that the polar caps were particularly error prone, as seen in \fig{fig:eb-maps-compare}. We excised these error prone regions, specifically masking portions of the maps with $|b|>30^{\circ}$ to find that the errors on the B mode spectra revert to having very similar structure as that of the E mode spectra, as seen in the bottom row of \fig{fig:eb-spectra_rad_cutoff}. Finally we modify the use of standard Healpix routines to mimic the local convolution E \& B construction (see \app{sec:convolution_err} for details). The results from this exercise also concur with our understanding of the discrepant behaviour for the B-mode spectrum.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Separating Stokes Q \& U maps corresponding to E \& B modes of polarization}
In this section we present the results of evaluating the local convolution given in \eq{eq:op_qu2equbqu}, to decompose the total Stokes Q and U parameters into those correspond to the E modes and B modes of polarization respectively. We reiterate that this procedure only relies on real space operation on the total Stokes Q and U maps and do not go through the process of constructing the E and B modes as is usually the case.
%
\begin{figure}[!t] 
\centering
\subfigure[E Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-healpix.pdf}}
\subfigure[E Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-2beta.pdf}}
\subfigure[E Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-diff.pdf}}
\subfigure[E U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-healpix.pdf}}
\subfigure[E U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-2beta.pdf}}
\subfigure[E U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-diff.pdf}}

@@

%
\begin{figure}[!t] 
\centering
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{\textit{Top:} The thick lines in the figure on left depicts the E mode spectra derived from pure E Stokes Q and U parameters derived for different radial cutoffs on the convolution, the thin lines depict the B mode spectra for the residual B modes.}
\label{fig:equ-bqu-spectra_rad_cutoff}
\end{figure}
%
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
%--------------------------------------------------------
%--------------------------------------------------------
